---
title: "Geospatial_Poisson"
author: "Prahlad"
date: "2023-12-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(sf)
library(tidyverse)
```

Load the burglaries data

```{r}
burglaries_2023 <- read_csv("../data/burglaries_2023.csv")
```
Load the census tracts data

```{r}
tracts <- read_sf("../data/DC/DC.shp")
```

spatial join the census tracts and burglaries

```{r}
burglaries_2023_geo <- st_as_sf(burglaries_2023 |> 
                                  drop_na(latitude),
                                coords = c('longitude', 'latitude'),
                                crs = st_crs(tracts)
                                )
```



```{r}
tracts |> 
  ggplot() +
  geom_sf() +
  geom_sf(data = burglaries_2023_geo, size = 01)
```
Some outlier in the data !!
```{r}
burglaries_2023 <- burglaries_2023 |>
  filter(latitude > 35)
```

```{r}
burglaries_2023_geo <- st_as_sf(burglaries_2023 |> 
                                  drop_na(latitude),
                                coords = c('longitude', 'latitude'),
                                crs = st_crs(tracts)
                                )
```

```{r}
tracts |> 
  ggplot() +
  geom_sf() +
  geom_sf(data = burglaries_2023_geo, size = 01)
```
Perform a spatial join


```{r}
burglaries_zip <- st_join(burglaries_2023_geo, tracts, join = st_within)
```


```{r}
census <- read_csv("../data/census.csv")
```
```{r}
census |>
  summarise(n_distinct(tract))
```



```{r}
burglaries_census_merged <- full_join(burglaries_zip, census, by = join_by(TRACTCE == tract))
```

Has the same number of entries as burglaries_zip. Used a left_join

Q. Does it contain all census tracts?
NO

WIth a full_join 

```{r}
burglaries_census_merged |>
  summarise(n_distinct(TRACTCE))
```
Still one observation short? 


Move on to part 2


Aggregate the data by census tract. **Warning:** each incident can appear multiple times if there are multiple victims, so be sure that you aren't double-counting any incidents.



Which census tract had the highest number of burglaries? Which census tract had the highest number of burglaries per 1000 residents?


```{r}
burglaries_census_merged |>
  group_by(TRACTCE) |>
  count() |>
  arrange(desc(n))
```


Tract 016000 has the highest number of burglaries in total.

```{r}
burglaries_census_merged |>
  group_by(TRACTCE) |>
  count() |>
  mutate(count_per_1000_residents = n / population)
```
```{r}
is_numeric_result <- burglaries_census_merged %>%
  summarise(is_numeric = is.numeric(population))
```

This returned true.

so why the error?
```{r}
burglaries_census_merged |>
  group_by(TRACTCE) |>
  count() |>
  mutate(count_per_1000_residents = n / burglaries_census_merged$population) |>
  arrange(desc(n))
```
```{r}
burglaries_census_merged |>
  group_by(TRACTCE) |>
  count() |>
  arrange(desc(n))
```


